R Tools for Observations, Receptors and Footprints (rtorf) for processing atmospheric observations in NOAA-GML

### 2025-07-23 paper/paper.Rmd
paper

The resulting `data.table` contains `r ncol(df)` columns and 
`r nrow(df)` observations. 
Furthermore, the size of `data.table` is `r format(object.size(df), unit = "Gb")`.
The data includes observations between `r range(df$year)[1]` and 
`r range(df$year)[2]`. Now we can define some parameters to filter our data,
like the year 2020 and spatially data below 8000 meters above sea level (masl) and
focused over north America. Nevertheless, it can be modified to any region.

@@

Now we have a `data.table` that contains `r ncol(df)` columns and 
`r nrow(df)` observations. The size of `data.table` is 
`r format(object.size(df), unit = "Gb")`. Sometimes the data can be alaready
filtered every 20 seconds or for a different period of time.
However, raw data can be availalb ein a second-by-second basis. 
Under this case we may need to aggregate datah. In this example, we can add a column
of time in format "POSIXct" and cut the seconds every 20 seconds.
Usually, aircraft observations every 1 second. Then,
We can simplify the data by calculating averages every 20 seconds.
We perform this task by cutting time very 20 seconds. Then,
we add a new column with the mandatory name `key_time`, which 
will be used to aggregate data with "POSIXct" class, but every 20 seconds.

@@

@@

Now we have the data processed and ready to be exported. `rtorf` includes 
a number of functions to save the data as tabulated format in text, csv and 
CSVY[^3] are csv files with a YAML header. This functions can be seen in the documentation.
In this last part of the manuscript we will show some visualizations. We included
a function named `obs_plot` which plots data in long format using R base functions.
Here we see data for the month of March 2020. This useful function allows to plot several
sites and prints the x-axis range.

@@

Finally, we show some vertical profiles for the months of January and March of 2020.
We can see how during March of 2020 methane concentrations are lower than
January. This may be due the implementation of Lockdowns [@espinosa2023covid].
A manuscript focused on the impact of COVID-19 on methane emissions
will be submitted soon.

@@

```{r, echo=FALSE, out.width="90%", fig.cap="a) Time series, b) Monthly observations by altitude", fig.show='hold',fig.align='center'}
knitr::include_graphics(c("paper_files/obsplot_ts.png",
                          "paper_files/obsplot_alt.png"))
```

>>>>

The resulting `data.table` contains `r ncol(df)` columns and 
`r nrow(df)` observations. 
Furthermore, the size of `data.table` is `r format(object.size(df), unit = "Mb")`.
The data includes observations between `r range(df$year)[1]` and 
`r range(df$year)[2]`. Now we can define some parameters to filter our data,
like the year 2020 and spatially data below 8000 meters above sea level (masl) and
focused over north America. Nevertheless, it can be modified to any region.

@@

Now we have a `data.table` that contains `r ncol(df)` columns and 
`r nrow(df)` observations. The size of `data.table` is 
`r format(object.size(df), unit = "Mb")`. Sometimes the data can be already
filtered every 20 seconds or for a different period of time.
However, raw data can be available in a second-by-second basis. 
Under this case we may need to aggregate data. In this example, we can add a column
of time in format "POSIXct" and cut the seconds every 20 seconds.
Usually, aircraft observations every 1 second. Then,
We can simplify the data by calculating averages every 20 seconds.
We perform this task by cutting time very 20 seconds. Then,
we add a new column with the mandatory name `key_time`, which 
will be used to aggregate data with "POSIXct" class, but every 20 seconds.

@@

Now we have the data processed and ready to be exported. `rtorf` includes 
a number of functions to save the data as tabulated format in text, csv and 
CSVY[^3] are csv files with a YAML header. These functions can be seen in the documentation.

@@

In this last part of the manuscript we will show some visualizations. We included
a function named `obs_plot` which plots data in long format using R base functions.
Here we see data for the month of March 2020. This useful function allows to plot several
sites and prints the x-axis range.

@@

Finally, we show some vertical profiles for the months of January and March of 2020.
"We can see that methane concentrations during March of 2020 are lower than in January.
This may be due the implementation of Lockdowns [@espinosa2023covid].
A manuscript focused on the impact of COVID-19 on methane emissions
will be submitted soon.

@@

```{r, echo=FALSE, out.width="90%", fig.cap=" Example visualizations generated by a) Time series of CH₄ concentrations for the IAGOS site on March 2, 2020. (b) Vertical profiles of CH₄ concentrations from January (labeled '1') and March (labeled '3') 2020, illustrating the package's ability to compare different time periods.", fig.show='hold',fig.align='center'}
knitr::include_graphics(c("paper_files/obsplot_ts.png",
                          "paper_files/obsplot_alt.png"))
```
